---
title: "Time Series Analysis project - forecast three-month and ten-year treasury
  maturity rates with VAR"
author: "Lin Xinyang.Monica"
date: "2023-08-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup,message=FALSE,warning=FALSE,eval=TRUE}
r = getOption("repos")
options(repos = c(CRAN = "https://cloud.r-project.org"))
```

## Packages

Let's load packages we will need to use for analysis

```{r,message=FALSE,warning=FALSE,eval=TRUE}
install.packages("xts")
install.packages("lmtest")
install.packages("ggplot2")
install.packages("zoo")
install.packages("tidyverse")
install.packages("dplyr")
install.packages("vars")
install.packages("knitr")
install.packages("kableExtra")
library(xts)
library(lmtest)
library(ggplot2)
library(zoo)
library(tidyverse)
library(dplyr)
library(vars)
library(knitr)
library(kableExtra)

source(file="D://Desktop//testdf.R")
```

## Import and process data

We will import data used for analysis and coerce it into "xts" dataset indexed by date

```{r}
r3m_r10y<-read.csv(file="D://Desktop//mrates.csv")
head(r3m_r10y,10)
tail(r3m_r10y,10)
class(r3m_r10y)
r3m_r10y$date<-as.Date(x=r3m_r10y$date,format="%Y-%m-%d")
r3m_r10y<-xts(x=r3m_r10y[,-1],order.by =r3m_r10y$date)
```

Data visualization

```{r}
plot(r3m_r10y[,1:2],
  type="l",
  col=c("red","blue"),
  major.ticks="years",
  grid.ticks.on="years",
  grid.ticks.lty=3,
  main="Three-month and ten-year treasury constant maturity rates"
)
```

They seem non-stationary and have a certain long-run relationship. But we need to perform formal tests to confirm this.

## Cointegration

Let's check whether r3m and r10y are integrated of the same order

```{r}
testdf(variable=r3m_r10y$r3m,max.augmentations = 3)
testdf(variable=r3m_r10y$r10y,max.augmentations = 3)
```

P-value of both data sets are much larger than 0.05, so we cannot reject the null hypothesis of DF test assuming non-stationarity.

Now let's difference r3m and r10y

```{r}
testdf(variable = diff(r3m_r10y$r3m),max.augmentations = 3)
testdf(variable = diff(r3m_r10y$r10y),max.augmentations = 3)
```

After the first differencing, both become stationary because p-value is much smaller than 0.05. Therefore they are both integrated of the order 1.

Let's run a linear regression.

```{r}
r3m_r10y_line<- lm(r3m_r10y$r10y~r3m_r10y$r3m,data=r3m_r10y)
summary(r3m_r10y_line)
```

r10m=0.78090\*r3m+2.39147+e_t

Don't forget to check whether there is autocorrelation in its residuals!

```{r}
testdf(variable = residuals(r3m_r10y_line),max.augmentations = 3)
```

High p-value suggests no autocorrelation. Therefore they are cointegrated

We could also perform granger casuality test:

```{r}
granger1<-grangertest(r10y~r3m,data=r3m_r10y,order=3)
granger1

granger2<-grangertest(r10y~r3m,data=r3m_r10y,order=5)
granger2

granger3<-grangertest(r3m~r10y,data=r3m_r10y,order=3)
granger3

granger4<-grangertest(r3m~r10y,data=r3m_r10y,order=5)
granger4
```
We could confirm that r10y granger causes r3m.

## ECM(Error Correction Mechanism)
```{r}
r3m_r10y$residual=lag.xts(x=resid(r3m_r10y_line),k=1)
r3m_d<-diff(x=r3m_r10y$r3m)
r10y_d<-diff(x=r3m_r10y$r10y)
r3m_r10y_ecm=lm(r10y_d~r3m_d+r3m_r10y$residual,data=r3m_r10y)
summary(r3m_r10y_ecm)
```
Intercept is insignificant, so let's remove it.
```{r}
r3m_r10y_ecm=lm(r10y_d~r3m_d+r3m_r10y$residual-1,data=r3m_r10y)
summary(r3m_r10y_ecm)
```
r10y_d=0.42128 * r3m_d -0.02669 * (r10m-0.78090*r3m-2.39147)+e_t.
```{r}
1/0.02669
```
It takes around 37.46 periods to clear errors

## Select VAR models
```{r}
r3m_r10y<-r3m_r10y[,-3]
VARselect(r3m_r10y,lag.max =8,season = 12)
```
Information criteria suggests VAR(7) and VAR(3).

 We could also visualize the result above:
```{r}
VARselect(r3m_r10y,lag.max =8,season = 12)%>%
  .$criteria %>%
  t() %>%
  as_tibble() %>%
  mutate(number_of_lags=1:nrow(.)) %>%
  kbl(digits=3) %>%
  kable_classic("striped",full_width=F)
```

Let's start with VAR(3):
```{r}
r3m_r10y.var3<-VAR(r3m_r10y,
                   p=3,
                   season=12)
summary(r3m_r10y.var3)
```

Most seasonal dummies are insignificant, so let's remove them.
```{r}
r3m_r10y.var3<-VAR(r3m_r10y,
                   p=3,
                   )
summary(r3m_r10y.var3)
```

Let's diagnose its residual
```{r}
plot(r3m_r10y.var3)
```
```{r}
serial.test(r3m_r10y.var3)
serial.test(r3m_r10y.var3,type=c("BG"))
```
Null hypothesis of no autocorrelation should be rejected based on very small p-value. So there is autocorrelation in the residual.

VAR(7):
```{r}
r3m_r10y.var7<-VAR(r3m_r10y,
                   p=7,
                   season = 12)
summary(r3m_r10y.var7)
```
Seasonal dummies are insignificant, so let's remove them.
```{r}
r3m_r10y.var7<-VAR(r3m_r10y,
                   p=7)
summary(r3m_r10y.var7)
```

Let's diagnoze its residual
```{r}
plot(r3m_r10y.var7)
```
```{r}
serial.test(r3m_r10y.var7)
serial.test(r3m_r10y.var7,type=c("BG")) 
```
Although there is autocorelation in its residual according to Portmanteau test, it passes BG test. So let's keep VAR(7).

VAR(6):
```{r}
r3m_r10y.var6<-VAR(r3m_r10y,p=6)
summary(r3m_r10y.var6)
```
```{r}
serial.test(r3m_r10y.var6)
serial.test(r3m_r10y.var6,type=c("BG"))
```
There is autocorrelation in it, so VAR(6) should not be considered.

VAR(9):
```{r}
r3m_r10y.var9<-VAR(r3m_r10y,p=9)
summary(r3m_r10y.var9)
```

```{r}
serial.test(r3m_r10y.var9)
serial.test(r3m_r10y.var9,type=c("BG"))
```
VAR(9) could also work since we cannot reject the null hypothesis assuming no autocorelation in Portmanteau test and BG test.

Let's use information criteria to see which one is better
```{r}
AIC(r3m_r10y.var9,r3m_r10y.var7)
BIC(r3m_r10y.var9,r3m_r10y.var7)
```

Let's use VAR(7)

## Impulse Response Functions
```{r}
plot(irf(r3m_r10y.var7,n.ahead = 40))
plot(fevd(r3m_r10y.var7,n.ahead = 40))
```

## Forecast 12-month periods ahead with VAR(7)
Let's cut the data in the last 12-month periods
```{r}
tail(r3m_r10y,13)
r3m_r10y.short<-r3m_r10y["/2019-03-01"]
tail(r3m_r10y.short,12)
```
Let's use VAR(7) to fit the shorten data set
```{r}
r3m_r10y.short.var7<-VAR(r3m_r10y.short,
                         p=7)
```

Let's perform forecasting
```{r}
r3m_r10y.forecast<-predict(r3m_r10y.short.var7,
                           n.ahead = 12,
                           ci=0.95)

r3m_r10y.forecast
```
Forecast data is stored in the column called "fcst"

forecast of three-month and ten-year treasury maturity rates
```{r}
r3m_r10y.forecast$fcst$r3m
r3m_r10y.forecast$fcst$r10y
```
Let's corcer them to "xts" and change the names of each column
```{r}
r3m_forecast<- xts(r3m_r10y.forecast$fcst$r3m[,-4],
                   tail(index(r3m_r10y),12))
names(r3m_forecast)<-c("r3m_forecast","r3m_lower","r3m_upper")


r10y_forecast<-xts(r3m_r10y.forecast$fcst$r10y[,-4],
                   tail(index(r3m_r10y),12) )
names(r10y_forecast)<-c("r10y_forecast","r10y_lower","r10y_upper")
```

Let's merge them with the original dataset.
```{r}
r3m_r10y_merged<-merge(r3m_r10y,
  r3m_forecast,
  r10y_forecast
)

r3m_r10y_merged
```

 Let's plot for r3m and r10y.
```{r}
plot(r3m_r10y_merged["2018/",c("r3m","r3m_forecast","r3m_lower","r3m_upper")],
     col=c("black","red","blue","blue"),
     major.ticks="years",
   grid.ticks.on="years",
   grid.ticks.lty=3,
   main="Real values of three-month maturity rates vs forecast"
)

plot(r3m_r10y_merged["2018/",c("r10y","r10y_forecast","r10y_lower","r10y_upper")],
     col=c("black","red","blue","blue"),
     major.ticks="years",
     grid.ticks.on="years",
     grid.ticks.lty=3,
     main="Real values of ten-year maturity rates vs forecast"
)
```

## Forecast Accuracy Measures
```{r}
r3m_r10y_merged$mae.r3m   <-  abs(r3m_r10y_merged$r3m - r3m_r10y_merged$r3m_forecast)
r3m_r10y_merged$mse.r3m   <-  (r3m_r10y_merged$r3m - r3m_r10y_merged$r3m_forecast) ^2
r3m_r10y_merged$mape.r3m  <-  abs((r3m_r10y_merged$r3m - r3m_r10y_merged$r3m_forecast)/r3m_r10y_merged$r3m)
r3m_r10y_merged$amape.r3m <-  abs((r3m_r10y_merged$r3m - r3m_r10y_merged$r3m_forecast) / 
                            (r3m_r10y_merged$r3m + r3m_r10y_merged$r3m_forecast))


r3m_r10y_merged$mae.r10y   <-  abs(r3m_r10y_merged$r10y - r3m_r10y_merged$r10y_forecast)
r3m_r10y_merged$mse.r10y  <-  (r3m_r10y_merged$r10y - r3m_r10y_merged$r10y_forecast) ^2
r3m_r10y_merged$mape.r10y  <-  abs((r3m_r10y_merged$r10y - r3m_r10y_merged$r10y_forecast)/r3m_r10y_merged$r10y)
r3m_r10y_merged$amape.r10y <-  abs((r3m_r10y_merged$r10y - r3m_r10y_merged$r10y_forecast) / 
                                    (r3m_r10y_merged$r10y + r3m_r10y_merged$r10y_forecast))

tail(r3m_r10y_merged,12)
dim(r3m_r10y_merged)
error<-colMeans(r3m_r10y_merged[,9:16],na.rm = TRUE)
error
```